Spatiotemporal models of species biomass split by maturity class

### 2019-08-02 analysis/VOCC/1-biomass-models.Rmd
minor tweaks to 1-biomass-models.rmd

# # Species run so far...
# species <- "Arrowtooth Flounder"
# species <- "Pacific Cod"
# species <- "Sablefish"
# species <- "Silvergray Rockfish"
# species <- "Lingcod"
# species <- "North Pacific Spiny Dogfish" # using all data for maturity thresholds
# species <- "Quillback Rockfish" # no maturity classes and WCHG not work
# species <- "Pacific Ocean Perch" 
species <- "Yelloweye Rockfish"
```

Region 
```{r}
# region <- "Both odd year surveys"
  region <- "West Coast Vancouver Island"
 region <- "West Coast Haida Gwaii"
```

@@

# covariates <- ""

  covariates <- "+muddy+any_rock"

@@

Run all subsequent code...
```{r global_options, include=FALSE}
covs <- gsub("\\+", "-", covariates)
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

@@

Run climate independent sdmTMB model
```{r}
  biomass <- readRDS(paste0(
    "data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"
    ))
  covars <- readRDS("data/event-covariates.rds")
  data <- dplyr::left_join(biomass, covars)

  # scale predictors before filtering to ensure mean and SD are global
  data <- data %>% mutate(raw_depth = depth, depth = log(raw_depth)) 
  data <- gfranges::scale_predictors(data, predictors = c(quo(depth)))
       # c(quo(log_depth), quo(mixed), quo(muddy), quo(sandy), quo(rocky), quo(any_rock)))
  data <- data %>% mutate(depth = raw_depth) %>% filter(ssid %in% model_ssid)

@@

    total_biomass <- sdmTMB::sdmTMB(data,
      dens_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime3 <- Sys.time()
    time3 <- round(starttime3 - endtime3)

@@

```{r}

@@

    nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
    nd <- nd_all %>% 
      filter(ssid %in% model_ssid) %>% 
      filter(year %in% unique(adult_biomass$data$year))
    nd <- na.omit(nd)
    nd$year <- as.integer(nd$year)

    adult_biomass <- readRDS(paste0("data/", spp, 
      "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
    ))
    adult_predictions <- predict(adult_biomass, newdata = nd)
    saveRDS(adult_predictions, file = paste0("data/", spp, 
      "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"
    ))

>>>>

# # Species run so far...
# species <- "Arrowtooth Flounder"
# species <- "Pacific Cod"
# species <- "Sablefish"
# species <- "Silvergray Rockfish"
 species <- "Lingcod"
# species <- "North Pacific Spiny Dogfish" # using all data for maturity thresholds
# species <- "Quillback Rockfish" # no maturity classes and WCHG not work
# species <- "Pacific Ocean Perch" 
# species <- "Yelloweye Rockfish"
```

Region 
```{r}
 region <- "Both odd year surveys"
region <- "West Coast Vancouver Island"
#  region <- "West Coast Haida Gwaii"
```

@@

covariates <- ""
covs <- gsub("\\+", "-", "+muddy+any_rock")
covariates <- "+muddy+any_rock"

@@

Run all subsequent code...
```{r global_options, include=FALSE}
# covs <- gsub("\\+", "-", covariates)
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

@@

Load and filter data 
```{r}
biomass <- readRDS(paste0(
    "data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"
    ))
  covars <- readRDS("data/event-covariates.rds")
  data <- dplyr::left_join(biomass, covars)

  # scale predictors before filtering to ensure mean and SD are global
  data <- data %>% mutate(raw_depth = depth, depth = log(raw_depth)) 
  data <- gfranges::scale_predictors(data, predictors = c(quo(depth)))
       # c(quo(log_depth), quo(mixed), quo(muddy), quo(sandy), quo(rocky), quo(any_rock)))
  data <- data %>% mutate(depth = raw_depth) %>% filter(ssid %in% model_ssid)
```

Run climate independent sdmTMB model
```{r}

@@

    total_biomass <- sdmTMB::sdmTMB(data,
      dens_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2 + trawled,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime3 <- Sys.time()
    time3 <- round(starttime3 - endtime3)

@@

```{r}
#covs <- gsub("\\+", "-", covariates)
fish <- readRDS(paste0("raw/bio-data-", spp, ""))

@@

    nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
    adult_biomass <- readRDS(paste0("data/", spp, 
      "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
    ))

    nd <- nd_all %>% 
      filter(ssid %in% model_ssid) %>% 
      filter(year %in% unique(adult_biomass$data$year))
    nd <- na.omit(nd)
    nd$year <- as.integer(nd$year)

    adult_predictions <- predict(adult_biomass, newdata = nd)
    saveRDS(adult_predictions, file = paste0("data/", spp, 
      "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"
    ))
